---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# populate

Draft for discussion.

This issue was closed : https://github.com/tidyverse/dplyr/issues/6547

Is this extension package useful ?

{populate} provides stricter wrappers around `dplyr::mutate()`:

* `populate()` makes sure the input's ptype is preserved
  * no column type change
  * no column addition
  * no column deletion (`mutate()` can delete with `NULL`)
  * cast by default, assert optionally
* `collate()` only creates new columns

Do we need more or should we keep it simple?

* `abrogate()` to remove columns ?
* `sublimate()` like `populate()` but allow ptype change (doesn't allow new col creation) ?
* `summarize()` variants ?

Is it better to use a subclass of data frame (as in linked github issue) so we can make safe versions of
everything including functions like `count()`, `unnest()`, joins ... 

We'd lose autocomplete of new args, maybe more confusing than having new verbs ?

## Installation

You can install the development version of populate like so:

``` r
devtools::install_github("cynkra/populate")
```

## Examples


```{r example, error = TRUE}
library(populate)

data <- tibble::tibble(
  a = letters[1:2],
  b = c(1,2), 
  c = factor(letters[1:2]), 
  d = as.Date(c("2022-01-01", "2022-01-02")), 
  e = vctrs::list_of(cars)
)

# can't create a column if it exists
collate(data, a = 1) 

# but we can create new columns
collate(data, ee = 1) 

# we can't create a new column with populate()
populate(data, ee = 1) 

 # can't cast double to character
populate(data, a = 1)

# casting integer to double
populate(data, b = 3:4) 

# doesn't work if `.strict` is `TRUE`
populate(data, b = 3:4, .strict = TRUE) 

# casting character to factor with allowed levels
populate(data, c = c("b", "b")) 

# can't cast because wrong levels
populate(data, c = c("b", "d")) 

# datetimes are casted to date
populate(data, d = lubridate::as_datetime(c("2022-01-01", "2022-01-02"))) 

 # characters can't be casted to date
populate(data, d = c("2022-01-01", "2022-01-02"))

# using list_of allowed us to prevent corrupting our data silently
populate(data, e = list(iris)) 

# and we don't have to bother with list_of anymore if we feed the right format
populate(data, e = list(head(cars))) 
```
